### 2017-05-09 paper2.tex
Add files via upload

\chapter{A climatology of urban surface heat islands derived from hemispherical radiometric surface temperatures}
\label{paper2}

@@

The temperature of the surface is integral in understanding, predicting, and modeling boundary-layer air temperature patterns, surface energy balances, and, in urban areas, has important implications for human thermal comfort and building energy usage. Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature (T\textsubscript{surf}). Integrated up to larger scales, urban areas tend to store more heat relative to non-built surroundings; traffic, HVAC systems, and industrial processes contribute to an additional anthropogenic heat input. These manifest in elevated T\textsubscript{surf} and T\textsubscript{air} in cities, a phenomenon termed the urban heat island effect (UHI). To foster a more complete understanding of the effect of urban areas on climates across a range of scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology. The proliferation of satellite and aerial thermal infrared (TIR) remote sensing has enabled spatially-extensive characterizations of surface climates at ever improving spatial and spectral resolutions. Such campaigns have elucidated urban T\textsubscript{surf} and surface urban heat island (sUHI) patterns globally at large spatial scales \cite{Peng2012,Zhao2014}. However, technological improvements in TIR remote sensing have yet to address three potential sources of error that arise when applied in urban areas:

@@

\subsection{Bias in thermal remote sensing}
Geometric biases in remote sensing of the urban surface are a result of its 3-dimensional, convoluted structure. Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing terrestrial radiation. The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms. Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry. This directional dependence of urban surface temperature is termed "effective thermal anisotropy" \citep{Voogt1998a}. Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface, leading to spatiotemporally variant directional biases in remote sensed urban T\textsubscript{surf} of up to 10 \si{\kelvin} \citep{Voogt1995}. In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \citep{Adderley2015}. However, the magnitude and diurnal and seasonal behaviors of this bias are dependent on sensor viewing geometry and unique site characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.). Thus, parameterization schemes to account for urban effective anisotropy are difficult to generalize across urban sites, sensor types, and sensor-surface geometries.

In addition to undersampling the urban surface, most thermal remote sensing platforms yield an instantaneous ‘snap-shot’ and cannot characterize time-continuous T\textsubscript{surf} patterns without sacrificing ground resolution. Temporal discontinuities in thermal remote sensing result in myriad potential sources of bias over a wide range of time scales. Aerial and satellite thermal remote sensing require clear sky conditions (clouds are opaque with respect to thermal infrared radiation). Hence, long term satellite characterizations of sUHI are biased towards conditions that maximize urban-rural contrasts in T\textsubscript{surf}. This results in an overestimation of “all-sky” sUHI. At diurnal scales, satellite overpass cycles rarely coincide with sUHI maximums and are not standard across cities or platforms. Thus, analysis of sUHI patterns and magnitudes across cities and instrument platforms is difficult. At shorter time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \citep{Christen2012}. Most thermal remote sensors provide instantaneous T\textsubscript{surf} (rather than temporally averaged) and are potentially contaminated by high-frequency microscale fluctuations in urban T\textsubscript{surf}. This is particularly salient in urban environments, where a large variety of fabric materials can produce significant spatial contrasts in thermal admittance (for instance, roof materials tend to have much lower thermal admittances than wall or road materials. Such large spatial contrasts in thermal admittance are generally not present in non-urban or rural environments). As a result, the magnitude of potential biases in remote sensed urban T\textsubscript{surf} from microscale fluctuations depends on the facet material types viewed by the sensor and can change with viewing direction and sensor FOV. The effect of this phenomenon on thermal remote sensing has not been extensively studied, however, the magnitude of microscale fluctuations in T\textsubscript{surf} is significant relative to a typical sUHI signal and thus constitutes a potentially large source of bias.

Both geometric and temporal shortcomings limit the representativity of traditional remote sensed evaluations of urban T\textsubscript{surf} and sUHI. The magnitude of these biases has not been extensively studied, particularly from a long term, climatological perspective. These biases lead us to ask: What is the nature of the true sUHI? How does it vary temporally and over a full range of weather conditions? To address these questions, this study presents the first time-continuous, climatological analysis of sUHI retrieved from radiometric hemispherical urban T\textsubscript{surf} retrieved via the correction method described in Chapter \ref{paper1}. These measures are used to assess the magnitude of and overcome geometric and temporal biases inherent in urban TIR remote sensing.

@@

Remote sensing of TIR is subject to atmospheric influence from gaseous and aerosol absorbers between the surface and the sensor. A pyrgeometer's broad waveband and wide FOV increase the potential for significant atmospheric influence on an observed TIR signal. These effects can result in differences of up to 8 \si{\kelvin} between the 'true' radiometric T\textsubscript{surf} and brightness T\textsubscript{surf} derived from remote sensed observations of upwelling TIR. Atmospheric effects are particularly important when comparing irradiances or remote sensed T\textsubscript{surf} across different study sites and times, as differences in surface geometry, instrument height, and ambient conditions can introduce large spatiotemporal contrasts in the magnitude of atmospheric influence. This is particularly salient for near-ground, broadband radiometers, as band by band spectral atmospheric transmittance can change significantly with path length, shown in Figure \ref{spectransheight}. As inter-site comparison is paramount in understanding the urban effect on T\textsubscript{surf} and is the basis of sUHI analysis, accurate atmospheric correction of remote sensed T\textsubscript{surf} is of prime importance. Thus, urban T\textsubscript{hem, r} in this study are derived using a multiple line-of-sight correction routine which accounts for the 3-dimensionality of the urban surface, changing atmospheric conditions, and spectrally non-uniform sensor response. For each time step, the method generates of modeled at-sensor irradiances for a range of potential T\textsubscript{hem, r} using profiles of measured T\textsubscript{air} and humidity. A corrected T\textsubscript{hem, r} for the target time step is then retrieved by matching the measured irradiance value to the closed modeled irradiance - T\textsubscript{hem, r} pairing. The work flow, described in Figure \ref{flow}, is repeated at 30-minute intervals over the eight month study period to retrieve a climatology of urban T\textsubscript{hem, r} for sUHI analysis.

First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \citep{Soux2004}. Using a digital building model, SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor. Path lengths are binned at 5\si{\degree} intervals and averaged over the azimuth angle. Finally, view factors are calculated for each 5\si{\degree} angular as each bin's proportion of the total view factor and normalized to unity.

@@

Although MODTRAN has been show to accurately model radiative transfer of broadband longwave radiation at urban scale path lengths \citep{Hoch2005, Hoch2007}, the added complexity inherent in simulating fluxes of radiation hemispherically over complex urban terrain prompted an independent evaluation of the correction method using observations of upwelling longwave radiation (measured from 2 \si{\meter}, 10 \si{\meter}, and 30 \si{\meter} above ground level) over a planar surface for a continuous 16-day period. In the evaluation, a modified version of the method uses T\textsubscript{hem, b} inferred from the lowest L\textsubscript{up} measurement along with profiles of T\textsubscript{air} and humidity to model irradiances at 10 \si{\meter} and 30 \si{\meter} at each time step. Modeled and measured irradiances match well, over wide range of mid-latitude atmospheric conditions, with results summarized in Table \ref{evalstats}.

By virtue of requiring nearly 500 radiative transfer simulations per time step and significant post-processing, the method described in \ref{method2} is computationally expensive. This prompted the development of a parameterization scheme to facilitate more widespread usage of urban T\textsubscript{hem, r} for climatological and comparative sUHI analysis. In the parametrization, to reduce computational time, radiative transfer simulations are run for a single view factor weighted average path length representing the average distance from the surface to the sensor as "seen" by the sensor over its FOV. This view factor weighted "hemispherical" path length can be calculated using a sensor view model (such as SUM) or approximated by hand using a digital building model (although care should be taken to ensure the sensor cosine response is accurately represented). Using the view factor weighted path length, a suite of radiative transfer simulations is run prior to correction to retrieve "hemispherical" broadband atmospheric transmittances for a range of potential humidities. Humidity-transmittance pairings are aggregated into a lookup table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance. Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) is inferred. Comparison of T\textsubscript{hem, r} derived from the correction method and the parameterization show good agreement, the results of which are included in Figure \ref{para_eval} and Table \ref{parastats}.

@@

This study uses data from two long-term climatological campaigns: the Basel Urban Boundary Layer Experiment (BUBBLE) conducted over December 2001 through July 2002 in Basel, Switzerland \citep{Rotach2005}, and the Environmental Prediction in Canadian Cities (EPiCC) (CITE?) campaign in Vancouver, Canada. Using the method described in section \ref{method2}, an eight month climatology of T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected as a part of the BUBBLE campaign. Urban data were observed from a tower located near city center at the "Basel Sperrstrasse" site. Rural reference data were observed at the "Lange Erlen" site approximately 6 \si{\kilo\meter} to the north east in the outskirts of the city of Basel.

@@

To supplement data from the BUBBLE campaign, radiation and meteorological variables observed as a part of the EPiCC project are used to derive a climatology of T\textsubscript{hem, r} for the year long period spanning October 2008 through September 2009 using the parameterization scheme. These data provide a second, long term, time continuous climatology of sUHI for comparison of seasonal and diurnal patterns, to understand the influence of urban and rural site selection and characteristics on sUHI, and to aid in developing a generalized model for sUHI development over multiple time scales.

@@

sUHI and canopy later UHI (clUHI) magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals. Rural T\textsubscript{hem, b} was calculated via

@@

To investigate the influence of rural site characteristics on Basel's sUHI, sUHI magnitudes are calculated using T\textsubscript{hem, b} inferred for a different rural site ("Village Neuf") over the summertime IOP. Observations at Village Neuf were intermittent and often discontinous. As such, a subset of nine partially-complete days was selected to represent T\textsubscript{hem, b} and sUHI over a range of cloud coverages, to investigate the influence of rural site characteristics and T\textsubscript{surf} regimes on sUHI. As sUHI develop, preference was given to days with continuois observations in early afternoon through night. For each of the nine days sUHI\textsubscript{max} was used to infer a normalized sUHI magnitude for both Lange Erlen and Village Neuf sites at 30- \si{\minute} intervals, to remove day-to and compare patterns of sUHI development for the two sites. Mean normalized sUHI was then calculated at each time step over the nine day period. A normalized T\textsubscript{hem, b} was also calculated (using maximum T\textsubscript{hem, b} for each day) to compare T\textsubscript{surf} regimes between the two sites.

@@

@@

Figure \ref{bx_suhi_compare} shows sUHI magnitudes calculated from urban T\textsubscript{comp}, T\textsubscript{plan}, and T\textsubscript{hem}. Compared to sUHI viewed from a complete representation of the urban surface (sUHI\textsubscript{comp}); nadir (sUHI\textsubscript{nadir}) and hemispherical (sUHI\textsubscript{hem}) views of the surface overestimate sUHI\textsubscript{comp} by day and underestimate sUHI\textsubscript{comp} by night. sUHI\textsubscript{nadir} from a nadir view shows the greatest diurnal range of magnitudes, particularly under clear sky 'satellite friendly' conditions.

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_iop_bubble}
	\caption{A comparison of sUHI magnitudes from complete (sUHI\textsubscript{comp}), hemispherical (sUHI\textsubscript{hem}), and nadir (sUHI\textsubscript{nadir}) remote sensed representations of the Sperrstrasse canyon over the IOP. Each plot includes mean sUHI, as well case days representing sUHI under clear sky and clouded sky conditions.}
	\label{bx_suhi_compare}
\end{figure}

@@

sUHI development is the result of urban modification to surface geometry and thermal and radiative properties, most notably sky-view-factor, impervious surface fraction, and thermal admittance (and by relation, urban-rural soil moisture contrasts). These differences alter urban heating/cooling rates relative to non-urban spaces and provide the genesis for diurnal sUHI development. In the early morning hours, rural surfaces respond quickly to insolation at low solar angles, and rural T\textsubscript{surf} rapidly increases. In contrast, shading by urban features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunrise. This is observed in Basel across all representations of urban T\textsubscript{surf} at similar magnitudes. Through the late morning hours, urban heating rates increase relative to rural rates as the solar angle increases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up. Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces. After the initial spike in rural heating in the early morning hours, urban heating rates overtake rural approximately three to four hours after sunrise and the sUHI slowly grows through midday to a peak at approximately two to three hours hours after solar noon. In the hours after UHI\textsubscript{max}, the solar angle decreases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to begin to cool.

In the afternoon hours, both urban and rural surfaces transition to cooling, this can lead to several different patterns in evening sUHI development and is strongly controlled by urban surface structure. Relative to non-urban spaces, cooling rates for wall and road facets lag slightly as a result of canyon radiation trapping, this is offset by increased cooling rates for low thermal admittance rooftop facets. These two factors control whether sUHI grows or shrinks in the hours after sUHI\textsubscript{max}. In urban areas with relatively unobstructed sky-view-factors or large rooftop plan area fractions, strong radiative cooling can lead to a sharp reduction in sUHI. The opposite is true of urban areas with restricted sky view factors and tall street canyons with large wall surface area fractions, where canyon radiative trapping reduces radiative cooling and sUHI is maintained through the evening. For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping. The dependence of sUHI development on canyon geometry is illustrated well by comparing nadir and complete representations of sUHI observed in Basel. A nadir view, by virtue of oversampling rapidly cooling rooftops, observes a strong evening decline in sUHI. This is roughly equivalent to sUHI calculated from an urban site with low canyon height to width ratio and a relatively unobstructed sky view factor. In contrast, by sampling vertical facets that cool more gradually, this decline is less evident in sUHI calculated from a complete view of Basel T\textsubscript{surf}. Thus, as expected, remote sensed sUHI analysis is also directionally dependent and subject to anisotropic effects when viewed from a narrow-FOV sensor. These effects not only affect the magnitude of remote sensed sUHI, but may also alter perceived diurnal patterns of sUHI.

After sunset, the unobstructed rural sky-view-factor allows for strong radiative cooling of the surface and rural cooling rates spike relative to urban, leading to a slight jump in sUHI. However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates. All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics. This is illustrated by two different patterns of post-sunset sUHI development observed for the two rural sites in Basel. Although similar in terms of surface coverage and site geometry, hilly terrain surrounding the Lange Erlen site results in cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days. This effect is not present at the Village Neuf site, where nocturnal cooling rates are likely driven primarily driven by radiative losses. The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site. Following site dependent effects on sUHI in the hours after sunset, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural.

@@

>>>>

\chapter{A climatology of sUHI derived from hemispherical radiometric surface temperatures}
\label{paper2}

@@

The temperature of the surface is integral in understanding, predicting, and modeling boundary-layer air temperature patterns, surface energy balances, and, in urban areas, has important implications for human thermal comfort and building energy usage. Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature (T\textsubscript{surf}). Integrated up to larger scales, urban areas tend to store more heat relative to non-built surroundings; traffic, HVAC systems, and industrial processes contribute to an additional anthropogenic heat input. In addition, a reduced pervious (and vegetative) cover in urban areas shifts convective fluxes towards sensible and away from latent. These manifest in elevated T\textsubscript{surf} and T\textsubscript{air} in cities, a phenomenon termed the urban heat island effect (UHI). To foster a more complete understanding of the effect of urban areas on climates across a range of scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology. The proliferation of satellite and aerial thermal infrared (TIR) remote sensing has enabled spatially-extensive characterizations of surface climates at ever improving spatial and spectral resolutions. Such campaigns have elucidated urban T\textsubscript{surf} and surface urban heat island (sUHI) patterns globally at large spatial scales \cite{Peng2012,Zhao2014}. However, technological improvements in TIR remote sensing have yet to address three potential sources of error that arise when applied in urban areas:

@@

\subsection{Bias in thermal remote sensing}
Geometric biases in remote sensing of the urban surface are a result of its 3-dimensional, convoluted structure. Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing longwave radiation. The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms. Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry. This directional dependence of urban surface temperature is termed "effective thermal anisotropy" \citep{Voogt1998a}. Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface, leading to spatiotemporally variant directional biases in remote sensed urban T\textsubscript{surf} of up to 10 to 12 \si{\kelvin} \citep{Voogt1995,Lagouarde2012}. In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \citep{Adderley2015}. However, the magnitude and diurnal and seasonal behaviors of this bias are dependent on sensor viewing geometry and unique site characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.). Thus, parameterization schemes to account for urban effective anisotropy are difficult to generalize across urban sites, sensor types, and sensor-surface geometries.

In addition to undersampling the urban surface, most thermal remote sensing platforms yield an instantaneous ‘snap-shot’ and cannot characterize time-continuous T\textsubscript{surf} patterns without sacrificing ground resolution. Temporal discontinuities in thermal remote sensing result in myriad potential sources of bias over a wide range of time scales. Aerial and satellite thermal remote sensing require clear sky conditions (clouds are opaque with respect to thermal infrared radiation). Hence, long term satellite characterizations of sUHI are biased towards conditions that maximize urban-rural contrasts in T\textsubscript{surf}. This results in an overestimation of “all-sky” sUHI. At diurnal scales, satellite overpass cycles rarely coincide with sUHI maximums and are not standard across cities or platforms. Thus, analysis of sUHI patterns and magnitudes across instrument platforms is difficult. At shorter time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \citep{Christen2012}. Most thermal remote sensors provide instantaneous T\textsubscript{surf} (rather than temporally averaged) and are potentially contaminated by high-frequency microscale fluctuations in urban T\textsubscript{surf}. This is particularly salient in urban environments, where a large variety of fabric materials can produce significant spatial contrasts in thermal admittance (for instance, roof materials tend to have much lower thermal admittances than wall or road materials. Such large spatial contrasts in thermal admittance are generally not present in non-urban or rural environments). As a result, the magnitude of potential biases in remote sensed urban T\textsubscript{surf} from microscale fluctuations depends on the facet material types viewed by the sensor and can change with viewing direction and sensor FOV. The effect of this phenomenon on thermal remote sensing has not been extensively studied, however, the magnitude of microscale fluctuations in T\textsubscript{surf} is significant relative to a typical sUHI signal and thus constitutes a potentially large source of bias.

Both geometric and temporal shortcomings limit the representivity of traditional remote sensed evaluations of urban T\textsubscript{surf} and sUHI. The magnitude of these biases has not been extensively studied, particularly from a long term, climatological perspective. These biases lead us to the question: What is the nature of the true sUHI? How does it vary temporally and over a full range of weather conditions? To address these questions, this study presents the first time-continuous, climatological analysis of sUHI retrieved from radiometric hemispherical urban T\textsubscript{surf} retrieved via the correction method described in Chapter \ref{paper1}. These measures are used to assess the magnitude of and overcome geometric and temporal biases inherent in urban TIR remote sensing.

@@

Remote sensing of TIR is subject to atmospheric influence from gaseous and aerosol absorbers between the surface and the sensor. A pyrgeometer's broad waveband and wide FOV increase the potential for significant atmospheric influence on an observed TIR signal. These effects can result in differences of up to 8 \si{\kelvin} between the 'true' radiometric T\textsubscript{surf} and brightness T\textsubscript{surf} derived from remote sensed observations of upwelling TIR. Atmospheric effects are particularly important when comparing irradiances or remote sensed T\textsubscript{surf} across different study sites and times, as differences in surface geometry, instrument height, and ambient conditions can introduce large spatiotemporal contrasts in the magnitude of atmospheric influence. This is particularly salient for near-ground, broadband radiometers, as band by band spectral atmospheric transmittance can change significantly with path length, shown in Figure \ref{spectransheight}. As inter-site comparison is paramount in understanding the urban effect on T\textsubscript{surf} and is the basis of sUHI analysis, accurate atmospheric correction of remote sensed T\textsubscript{surf} is of prime importance. Thus, urban T\textsubscript{hem, r} in this study are derived using a multiple line-of-sight correction routine which accounts for the 3-dimensionality of the urban surface, changing atmospheric conditions, and spectrally non-uniform sensor response. For each time step, the method generates a LUT of modeled at-sensor irradiances for a range of potential T\textsubscript{hem, r} using profiles of measured T\textsubscript{air} and humidity. A corrected T\textsubscript{hem, r} for the target time step is then retrieved by matching the measured irradiance value to the closed modeled irradiance - T\textsubscript{hem, r} pairing. The work flow, described in Figure \ref{flow}, is repeated at 30-minute intervals over the eight month study period to retrieve a climatology of urban T\textsubscript{hem, r} for sUHI analysis.

First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \citep{Soux2004}. Using a digital building model (DBM), SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor. Path lengths are binned at 5\si{\degree} intervals and averaged over the azimuth angle. Finally, view factors are calculated for each 5\si{\degree} angular as each bin's proportion of the total view factor and normalized to unity.

@@

Although MODTRAN has been show to accurately model radiative transfer of broadband longwave radiation at urban canyon scale path lengths \citep{Hoch2005, Hoch2007}, the added complexity inherent in simulating fluxes of radiation hemispherically over complex urban terrain prompted an independent evaluation of the correction method using observations of upwelling longwave radiation (measured from 2 \si{\meter}, 10 \si{\meter}, and 30 \si{\meter} above ground level) over a simple, flat surface for a continuous 16-day period. In the evaluation, a modified version of the method uses T\textsubscript{hem, b} inferred from the lowest L\textsubscript{up} measurement along with profiles of T\textsubscript{air} and humidity to model irradiances at 10 \si{\meter} and 30 \si{\meter} at each time step. Modeled and measured irradiances match well, over wide range of mid-latitude atmospheric conditions, with results summarized in Table \ref{evalstats}.

By virtue of requiring nearly 500 radiative transfer simulations per time step and significant post-processing, the method described in \ref{method2} is computationally expensive. This prompted the development of a parameterization scheme to facilitate more widespread usage of urban T\textsubscript{hem, r} for climatological and comparative sUHI analysis. In the parametrization, to reduce computational time, radiative transfer simulations are run for a single view factor weighted average path length representing the average distance from the surface to the sensor as "seen" by the sensor over its FOV. This view factor weighted "hemispherical" path length can be calculated using a sensor view model (such as SUM) or approximated by hand using a digital building model (although care should be taken to ensure the sensor cosine response is accurately represented). Using the view factor weighted path length, a suite of radiative transfer simulations is run prior to correction to retrieve "hemispherical" broadband atmospheric transmittances for a range of potential humidities. Humidity-transmittance pairings are aggregated into a look-up table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance. Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) is inferred. Comparison of T\textsubscript{hem, r} derived from the correction method and the parameterization show good agreement, the results of which are included in Figure \ref{para_eval} and Table \ref{parastats}.

@@

This study uses data from two long-term climatological campaigns: the Basel Urban Boundary Layer Experiment (BUBBLE) conducted over December 2001 through July 2002 in Basel, Switzerland \citep{Rotach2005}, and the Environmental Prediction in Canadian Cities (EPiCC) \citep{Voogt2007} campaign in Vancouver, Canada. Using the method described in section \ref{method2}, an eight month climatology of T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected as a part of the BUBBLE campaign. Urban data were observed from a tower located near city center at the "Basel Sperrstrasse" site. Rural reference data were observed at the "Lange Erlen" site approximately 6 \si{\kilo\meter} to the north east in the outskirts of the city of Basel.

@@

To supplement data from the BUBBLE campaign, radiation and meteorological variables observed as a part of the EPiCC project are used to derive a climatology of T\textsubscript{hem, r} for the year long period spanning October 2008 through September 2009 using the parameterization scheme. EPiCC urban and rural data were collected from an urban/suburban site in central Vancouver ("Sunset Tower") and from a rural site approximately 25 \si{\kilo\meter} to the SSW ("Westham Island"). These data provide a second, long term, time continuous climatology of sUHI for comparison of seasonal and diurnal patterns, to understand the influence of urban and rural site selection and characteristics on sUHI, and to aid in developing a generalized model for sUHI development over multiple time scales.

@@

sUHI and clUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals. Rural T\textsubscript{hem, b} was calculated an expanded version of equation \ref{stefb1} to remove emissivity effects,

@@

To investigate the influence of rural site characteristics on Basel's sUHI, sUHI magnitudes are calculated using T\textsubscript{hem, b} inferred for a different rural site ("Village Neuf") over the summertime IOP. Observations at Village Neuf were intermittent and often discontinuous As such, a subset of nine partially-complete days was selected to represent T\textsubscript{hem, b} and sUHI over a range of cloud coverages, to investigate the influence of rural site characteristics and T\textsubscript{surf} regimes on sUHI. As sUHI develop, preference was given to days with continuous observations from early afternoon through the night. For each of the nine days sUHI\textsubscript{max} was used to infer a normalized sUHI magnitude for both Lange Erlen and Village Neuf sites at 30- \si{\minute} intervals to compare patterns of sUHI development for the two sites. Mean normalized sUHI was then calculated at each time step over the nine day period. A normalized T\textsubscript{hem, b} was also calculated (using maximum T\textsubscript{hem, b} for each day) to compare T\textsubscript{surf} regimes between the two sites.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_season_vanc}
	\caption{Hemispherical sUHI magnitudes for the year-long EPiCC climatology (top) and binned for winter (middle) and summer (bottom) seasons. Grey shading indicates nighttime hours averaged over the bin interval.}
	\label{bxsUHI_season_vanc}
\end{figure}

@@

Figure \ref{bx_suhi_compare} shows sUHI magnitudes calculated from urban T\textsubscript{comp}, T\textsubscript{plan}, and T\textsubscript{hem}. Compared to sUHI viewed from a complete representation of the urban surface (sUHI\textsubscript{comp}); nadir (sUHI\textsubscript{nadir}) and hemispherical (sUHI\textsubscript{hem}) views of the surface overestimate sUHI\textsubscript{comp} by day and underestimate sUHI\textsubscript{comp} by night. sUHI\textsubscript{nadir} shows the greatest diurnal range of magnitudes, particularly under clear sky 'satellite friendly' conditions.

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_iop_bubble}
	\caption{A comparison of sUHI magnitudes from complete (sUHI\textsubscript{comp}), hemispherical (sUHI\textsubscript{hem}), and nadir (sUHI\textsubscript{nadir}) remote sensed representations of the Sperrstrasse canyon over the IOP. Each plot includes mean sUHI, as well as case days representing sUHI under clear sky and clouded sky conditions.}
	\label{bx_suhi_compare}
\end{figure}

@@

sUHI development is the result of urban modification to surface geometry and thermal and radiative properties, most notably sky-view-factor, impervious surface fraction, and thermal admittance (and by relation, urban-rural soil moisture contrasts). These differences alter urban heating/cooling rates relative to non-urban spaces and provide the genesis for diurnal sUHI development. In the early morning hours, rural surfaces respond quickly to insolation at low solar angles, and rural T\textsubscript{surf} rapidly increases. In contrast, shading by urban features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunrise. This is observed in Basel across all representations of urban T\textsubscript{surf}. Through the late morning hours, urban heating rates increase relative to rural rates as the solar zenith angle decreases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up. Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces. After the initial spike in rural heating in the early morning hours, urban heating rates overtake rural approximately three to four hours after sunrise and the sUHI slowly grows through midday to a peak at approximately two to three hours hours after solar noon. In the hours after sUHI\textsubscript{max}, the solar zenith angle increases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to begin to cool.

In the afternoon hours, both urban and rural surfaces transition to cooling; this can lead to several different patterns in evening sUHI development and is strongly controlled by urban surface structure. Relative to non-urban spaces, cooling rates for wall and road facets lag slightly as a result of canyon radiation trapping, this is offset by increased cooling rates for low thermal admittance rooftop facets. These two factors control whether sUHI grows or shrinks in the hours after sUHI\textsubscript{max}. In urban areas with relatively unobstructed sky-view-factors or large rooftop plan area fractions, strong radiative cooling can lead to a sharp reduction in sUHI. The opposite is true of urban areas with restricted sky view factors and tall street canyons with large wall surface area fractions, where canyon radiative trapping reduces radiative cooling and sUHI is maintained through the evening. For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping. The dependence of sUHI development on canyon geometry is illustrated well by comparing nadir and complete representations of sUHI observed in Basel. A nadir view, by virtue of oversampling rapidly cooling rooftops, observes a strong evening decline in sUHI. In contrast, by sampling vertical facets that cool more gradually, this decline is less evident in sUHI calculated from a complete view of Basel T\textsubscript{surf}. Thus, as expected, remote sensed sUHI analysis is also directionally dependent and subject to anisotropic effects when viewed from a narrow-FOV sensor. These effects not only affect the magnitude of remote sensed sUHI, but may also alter perceived diurnal patterns of sUHI.

After sunset, the unobstructed rural sky-view-factor allows for strong radiative cooling of the surface and rural cooling rates spike relative to urban, leading to a slight jump in sUHI. However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates. All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics, similar to observations of variations in clUHI based on rural siting in \citep{Runnalls2000}. This is illustrated by two different patterns of post-sunset sUHI development observed for the two rural sites in Basel. Although similar in terms of surface coverage and site geometry, hilly terrain surrounding the Lange Erlen site results in suspected cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days. This effect is not present at the Village Neuf site, where nocturnal cooling rates are likely driven primarily driven by radiative losses. The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site. Following site dependent effects on sUHI in the hours after sunset, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural.

@@

Results of sensor placement sensitivity testing in Section \ref{sensitivity} show that T\textsubscript{hem, r} is dependent on sensor position relative to canyon features. These effects are translated into sUHI\textsubscript{hem}. In general, when increasing sUHI magnitudes.
